---
title: "Rethinking Chapter 4"
output: html_notebook
---
- Linear regression is the geocentric model of applied statistics.

# Normal by addition
R-Code 4.1
```{r}
pos <- replicate(1000, sum( runif(16,-1,1)))
#adjust plot margins
hist(pos)
plot(density(pos))
```
# Normal by multiplication
- Multiplying small numbers is approximately the same as addition.
- The interacting growth deviations, as long as they are sufficiently small, converge to a Gaussian distribution.


R-Code 4.2
```{r}
# Sample a random growth rate 
# This code samples 12 random numbers between 1 and 1.1.
prod (1 + runif(12,0,0.1))

# And this code generates 10.000 of these numbers
growth <-replicate(10000,prod(1+runif(12,0,0.01)))
library(rethinking)
dens( growth,norm.comp=TRUE)
```
# Normal by log-multiplication
- Multiplicative interactions of large deviations can produce Gaussian distributions, once we measure the outcomes on the log scale.

R-Code 4.5
```{r}
log.big <-replicate(10000,log(prod(1+runif(12,0,0.5))))
dens(log.big,norm.comp=TRUE)
```