---
title: "Rethinking Chapter 4"
output: html_notebook
---
- Linear regression is the geocentric model of applied statistics.

# Normal by addition
R-Code 4.1
```{r}
pos <- replicate(1000, sum( runif(16,-1,1)))
#adjust plot margins
hist(pos)
plot(density(pos))
```
# Normal by multiplication
- Multiplying small numbers is approximately the same as addition.
- The interacting growth deviations, as long as they are sufficiently small, converge to a Gaussian distribution.


R-Code 4.2
```{r}
# Sample a random growth rate 
# This code samples 12 random numbers between 1 and 1.1.
prod (1 + runif(12,0,0.1))

# And this code generates 10.000 of these numbers
growth <-replicate(10000,prod(1+runif(12,0,0.01)))
library(rethinking)
dens( growth,norm.comp=TRUE)
```
# Normal by log-multiplication
- Multiplicative interactions of large deviations can produce Gaussian distributions, once we measure the outcomes on the log scale.

R-Code 4.5
```{r}
log.big <-replicate(10000,log(prod(1+runif(12,0,0.5))))
dens(log.big,norm.comp=TRUE)
```
# From model definition to Bayes' theorem
R-Code 4.6
```{r}
w <- 6; n <- 9;
p_grid <- seq(from=0,to=1,length.out=100)
posterior <- dbinom(w,n,p_grid)*dunif(p_grid,0,1)
posterior <- posterior/sum(posterior)
```

# Gaussian model of height
## Get the test data
R-code 4.7
```{r}
library(rethinking)
data(Howell1)
d <- Howell1

str(d) # Inspects the structure of the data frame
precis(d,hist=FALSE)

d$height # Extracts the height as R vector

d2 <- d[ d$age >= 18, ] # Extract individuals of age 18 or greater

precis(d2,hist=FALSE)

dens(d$height) # Plot the distribution of heights

curve(dnorm(x, 187, 20), from=100, to=250) # Plot prior for mu
curve(dunif(x, 0, 50), from =-10, to=60) # Plot prior for sigma

# And now sample from prior to simulate heights
sample_mu <- rnorm(1e4, 187, 20)
sample_sigma <- runif(1e4, 0, 50)
prior_h <- rnorm(1e4, sample_mu, sample_sigma)
dens(prior_h)
```
# Grid approximation of the posterior
R code 4.16
```{r}
# Some crazy code
mu.list <-seq(from=150,to=160,length.out=100)
sigma.list <-seq(from=7,to=9,length.out=100)
post <-expand.grid(mu=mu.list,sigma=sigma.list)
post$LL <-sapply(1:nrow(post),function(i)sum(
dnorm( d2$height,post$mu[i],post$sigma[i],log=TRUE)))
post$prod <-post$LL+dnorm(post$mu,187,20,TRUE)+
dunif( post$sigma,0,50,TRUE)
post$prob <-exp(post$prod-max(post$prod))

contour_xyz(post$mu, post$sigma, post$prob)

image_xyz(post$mu, post$sigma, post$prob)

# Sampling from the posterior
# Note how we first randomly sample row numbers in post in proportion to the values in post$prob and then pull out the parameter values on those randomly sampled rows.

sample.rows <-sample(1:nrow(post),size=1e4,replace=TRUE,
prob=post$prob )
sample.mu <-post$mu[sample.rows]
sample.sigma <-post$sigma[sample.rows]

plot(sample.mu,sample.sigma,cex=0.5,pch=16,col=col.alpha(rangi2,0.1))

dens(sample.mu)
dens(sample.sigma)

PI(sample.mu)
PI(sample.sigma)
```
```{r}
# Now with only 20 of the heights from the height data

d3 <-sample(d2$height,size=20)

mu.list <-seq(from=150,to=170,length.out=200)
sigma.list <-seq(from=4,to=20,length.out=200)
post2 <-expand.grid(mu=mu.list,sigma=sigma.list)
post2$LL <-sapply(1:nrow(post2),function(i)
sum( dnorm(d3,mean=post2$mu[i],sd=post2$sigma[i],
log=TRUE )))
post2$prod <-post2$LL+dnorm(post2$mu,178,20,TRUE)+
dunif( post2$sigma,0,50,TRUE)
post2$prob <-exp(post2$prod-max(post2$prod))
sample2.rows <-sample(1:nrow(post2),size=1e4,replace=TRUE,
prob=post2$prob )
sample2.mu <-post2$mu[sample2.rows]
sample2.sigma <-post2$sigma[sample2.rows]

plot( sample2.mu,sample2.sigma,cex=0.5,col=col.alpha(rangi2,0.1),
      xlab="mu" ,ylab="sigma",pch=16)

dens(sample2.sigma,norm.comp=TRUE)

```

# Quadratic Approximation

```{r}
# Load data
library(rethinking)
data(Howell1)
d <-Howell1
d2 <-d[d$age>=18,]

# Define the model
flist <-alist(
height ~dnorm(mu,sigma),
mu ~dnorm(187,20),
sigma ~dunif(0,50)
)

# Fit the model to the data
m4.1 <-quap(flist,data=d2)

# Take a look at the posterior distribution
precis(m4.1)

# Now again, but with a more informative prior for mu
m4.2 <-quap(
  alist(
    height ~dnorm(mu,sigma),
    mu ~dnorm(187,0.1),
    sigma ~dunif(0,50)
    ) ,data=d2)
precis( m4.2)

# Show covariance between parameters (since we are dealing with a multi-dimensional Gaussian distribution)
vcov(m4.1)

# Decomposed into a vector of variances and a correlation matrix
diag(vcov(m4.1))
cov2cor(vcov(m4.1))

# Sample vectors of values
post <- extract.samples(m4.1, n=1e4)

# Show some of the samples
head(post)
precis(post)
```

